---
title: "Insurance_Claims_Analysis"
author: "Sai Varadharajan"
date: "8/13/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Loading in Libraries
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
```
### Reading in Dataset

```{r}
df_insurance <- read_csv("Insurance_factor_identification.csv")
```

```{r}
str(df_insurance)
```
## Analysis Task 1: The committee is interested to know each field of the data collected through descriptive analysis to gain basic insights into the data set and to prepare for further analysis. 

```{r}
summary(df_insurance)
```

## Data Preprocessing

Checking for Missing Values:
```{r}
sum(is.na(df_insurance)) # no missing values
```
Checking ranges of each column to make sure values make sense:

 Kilometers- Values should be between 1 and 5, 
 Zone: There are supposed to be 7 zones, from 1-7
 Bonus: Min value should be 1, since it is calculated by number of years since last claim + 1
 Make: There are 9 categories for Make
 Insured- Number of insured policy years. Seems to be positively Skewed.
 Claims- Seems to be positively Skewed.
 Payments: Seems to be positively Skewed.

factorizing necessary columns: Zone, Make
I might change the Km to a factor later since the values represent a range of distances traveled. I'll leave it as it is for now.

```{r}
df_processed_insurance <- df_insurance %>%
  mutate(
    Zone=as.factor(Zone),
    Make=as.factor(Make),
    Bonus=as.factor(Bonus),
    Kilometres=as.factor(Kilometres)
  )
str(df_processed_insurance)
```
Handling Outliers:
Insured, Claims, Payments

Viewing Distributions:
```{r}

q1 <- quantile(df_processed_insurance$Insured, 0.25)
median <- median(df_processed_insurance$Insured)
q3 <- quantile(df_processed_insurance$Insured, 0.75)


ggplot(df_processed_insurance) +
  geom_boxplot(aes(x = "", y = Insured)) +
  labs(title = "Box Plot of Insured",x=" Insurance Values") +
  scale_y_continuous(limits = c(0, 1500))+
geom_text(aes(x = 1, y = q1, label = paste("Q1:", q1)),
            vjust = -0.5, hjust = 0.5, color = "blue") +
  geom_text(aes(x = 1, y = median, label = paste("Median:", median)),
            vjust = -0.5, hjust = 1.5, color = "red") +
  geom_text(aes(x = 1, y = q3, label = paste("Q3:", q3)),
            vjust = 2.5, hjust = 0.5, color = "blue")

```
Viewing insured by Make

```{r}
make_vs_insured <- ggplot(df_processed_insurance)+
  geom_boxplot(aes(x=Make,y=Insured),fill="orange")

make_vs_insured
```

It seems like the extreme values are coming mostly from the ninth make, which is the Others Category






```{r}
ggplot(df_processed_insurance) +
  geom_histogram(aes(x = Claims), binwidth = 30, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Claims")
```
There seems to numerous claim counts past the interquartile range of 1-21, indicating significant variability and dispersion. These outliers seem valid and could possibly lead to insights, so I will keep them as it is without applying log transformations.

```{r}
make_vs_claims <- ggplot(df_processed_insurance) +
  geom_boxplot(aes(x = Make, y=Claims),fill="yellow") +
  labs(title = "Number of Claims per Make", y="Count of Claims")
make_vs_claims
```
It seems like the larger values for claims come from the "Others" section, aka Make 9. For the most part, claims are well below 500 for the other Makes.



```{r}
make_vs_payment <- ggplot(df_processed_insurance) +
  geom_boxplot(aes(x = Make, y=Payment,fill="red")) +
  labs(title = "Payments made for each Make", y="Swedish Krona Currency")
make_vs_payment
```
It makes sense why Make 9 has such a wide dispersion, since it represents the "Other" category. The dispersion and variability is fairly low for categories 2-8.



As we can see, most of the variability and extreme values comes from the "Others" Make category, Category 9
```{r}
# plotting Make against bonus as well

make_vs_bonus <- ggplot(df_processed_insurance) +
  geom_boxplot(aes(x = Make, y=Bonus,fill="pink")) +
  labs(title = "Make vs Bonus", y="Bonus")

(make_vs_claims + make_vs_payment) /(make_vs_insured + make_vs_bonus)

```
It seems like Category 9 does not have any strange outliers when it comes to the Bonus section, strangely.

### Distributions of Distance Traveled, Origin, and Make





```{r}
bar_make <- ggplot(df_processed_insurance)+
  geom_bar(aes(x=Make,fill=Make)) +
  labs(title="Number of each Make")

 make_by_zone <- ggplot(df_processed_insurance, aes(x = Make, fill = as.factor(Zone))) +
  geom_bar(position = "dodge") +
  labs(title = "Make By Zone") +
  theme_minimal()

make_vs_km <- ggplot(df_processed_insurance, aes(x = Make, y = Kilometres,fill=Make)) +
  geom_boxplot() +
  labs(title = "Make vs Kilometers Traveled") +
  theme_minimal()

bar_make +  make_by_zone /make_vs_km
```

The dataset seems to have a roughly equal number of each Make. This is interesting since the Ninth make, Others, is responsible for a disproportionally large amount of variance and outliers. Make 9 seems to have the same dispersion in Kilometers as well as zone  as compared to the other Makes, which is strange since it has such high variability and outliers in terms of Claims, Payments, and Insured.


Using hypotheseis testing to check if the observed differences between Makes 1-8 and 9 are due to chance or are statistically significant:

Insured Variable:

Null Hypothesis, H0- There is so significant difference between Insured status between Make 9 and the other makes
Alternative Hypothesis H1- There is a signficant difference between Insured status between the two groups.
Significance level: 0.05

```{r}
make9_data <- df_processed_insurance %>%
filter(Make == 9)

other_makes_data <- df_processed_insurance %>%
  filter(Make != 9)

# Perform independent samples t-test
ttest_result <- t.test(make9_data$Insured, other_makes_data$Insured)

# Print the t-test result
print(ttest_result)

```
```{r}
str(colnames(df_processed_insurance))

for (col in colnames(df_processed_insurance)) {
  print(col)
}
```


```{r}
make9_data <- df_insurance %>% # using original dataset because every column in original is numeric
filter(Make == 9)

other_makes_data <- df_insurance %>%
  filter(Make != 9)

# Perform independent samples t-test
for (col in colnames(df_insurance)) {
  cat("Checking column:", col, "\n")
  ttest_result <- t.test(make9_data[[col]], other_makes_data[[col]])
  # Print the t-test result
  print(ttest_result)
}




```
It seems like Make 9 is statistically significant in terms of Number of Claims, Payment, and Insured but not in terms of Kilometers traveled and Bonuses.

This could indicate that this category has specific characteristics that lead to higher claims, payments, and insured amounts. This might be due to certain features of the vehicles in this category, driving patterns, or other factors unique to "Make 9."

## Analysis Task 2: The total value of payment by an insurance company is an important factor to be monitored. So the committee has decided to find whether this payment is related to the number of claims and the number of insured policy years. They also want to visualize the results for better understanding. 

My approach:

Using regression lines and plotting Payment vs Insured and Claims and checking the regression line. I was also considering doing it separately for models 1-8 and model 9. 



Payment vs Insured

```{r}
ggplot(df_processed_insurance,aes(x=Payment,y=Insured)) +
  geom_point() +
  geom_smooth(method="lm")
```
R squared value:
```{r}
lm_model <- lm(Insured~ Payment, data = df_processed_insurance)
rsquared <- summary(lm_model)$r.squared
rsquared
```


Payment vs Claims


```{r}
ggplot(df_processed_insurance,aes(x=Payment,y=Claims)) +
  geom_point() +
  geom_smooth(method="lm")
```

R squared value:
```{r}
lm_model <- lm(Claims ~ Payment, df_processed_insurance)
summary(lm_model)$r.squared
```


As we can see, Payment is strongly positively correlated, with an R Squared value of 0.87 with both Insurance and Claims.



### Analysis Task 3
the committee wants to figure out the reasons for insurance payment increase and decrease. So they have decided to find whether distance, location, bonus, make, and insured amount or claims are affecting the payment or all or some of these are affecting it. 


Approach: I'll use multiple linear regression and check each independent variable for statistical significance, i.e. if their p-value is less than 0.05.


```{r}
multiplelm <- lm(Payment ~ .,df_insurance)
summary(multiplelm)
```
We can see that Kilometers, Zome, Insured and Claims are statistically significant while Bonus and Make are not, implying that the former affect Payment while the latter do not.

### Analysis Task 4:

The insurance company is planning to establish a new branch office, so they are interested to find at what location, kilometre, and bonus level their insured amount, claims, and payment gets increased. 

Approach : Group by location, km, bonus level and summarize mean amount, claims, and payment


```{r}
agg_data <- df_processed_insurance %>%
  group_by(Kilometres, Zone, Bonus) %>%
  summarise(
    mean_insured= mean(Insured),
    mean_claims=mean(Claims),
    mean_payment=mean(Payment)
    
  )
agg_data
```

```{r}
summary(agg_data)
```


```{r}
ordered_agg_data <- agg_data[order(-agg_data$mean_insured, -agg_data$mean_claims, -agg_data$mean_payment), ]
ordered_agg_data

```

Based on this table, I'd recommend Zone 4, which has the top 3 highest mean Insured , claims amd payment values.


### Analysis Task 5:

The committee wants to understand what affects their claim rates so as to decide the right premiums for a certain set of situations. Hence, they need to find whether the insured amount, zone, kilometre, bonus, or make affects the claim rates and to what extent. 

Approach: I will check the p values for statistical significance and check the coefficients for magnitude of effect.



```{r}
lm_claimrates <- lm(Claims ~Insured + Zone + Kilometres + Bonus, df_insurance)
summary(lm_claimrates)
```



We can see that the Insured amount, zone, kilometre, bonus all are statistically significant, meaning that they affect the Claims rate. Zone seems to have the strongest magnitude of effect at -6, followed by Bonus at -4 and Kilometers at -3, and then by Insured at 0.03. 

